Response to reviewer comments ELE-00738-2016

### 2016-12-09 paper/EL-reviewer-comments-1.Rmd
Theoretical framework

---
title: "Response to reviewer comments ELE-00738-2016"
date: "21 November 2016"
output: pdf_document
bibliography: references.bib
---

@@

We are grateful for the opportunity to resubmit a revised manuscript for your consideration. We have endeavored to incorporate all reviewers' feedback into the manuscript and we feel the attached draft constitutes an important improvement from the previous version. We were impressed by the quality of the reviews; they all were particularly thoughtful, clear, and constructive. It is evident that all three anonymous reviewers have given a fair and careful consideration to our manuscript. Therefore, we would like to especially thank them---not with the customary platitude usual in response letters, but with sincere gratitude for their valuable feedback.

The reviewers identified major limitations in our manuscript. Specifically with regards to the terminology, the underlying  structural control, and the lack of detail in the presentation of the empirical networks. In the current version, we have thoroughly addressed all these concerns and improved the robustness of the analyisis and the clarity of the concepts. Detailed responses to each comment, one by one, are provided below.

@@

    *Accepted; changed. We revised the text to better reflect the contribution of @Tylianakis2010 at the same time we added a reference to a more recent paper by @Kaiser-Bunbury2015 who explicitly deals with oportunities, limitations and implications of applying network metrics in conservation science.*

- Lines 95-97: "The number of nodes necessary to fully control a complex network can be calculated by counting the number of unmatched nodes in the network's maximum matching (Liu et al. 97 2011)." This statement must be contextualized by the assumed nature of the dynamical relationships in this framework. It therefore seems to me that this statement should be preceded by a discussion of those details (indeed, the authors argue that this is necessary in the Introduction, on lines 48-49).

- Furthermore, the authors need to discuss the appropriateness of this framework to ecological systems. The input control signals that drive even relatively simple networks to specified states can be very sensitive to initial conditions and complicated (indeed, they usually aren't calculated at all). How can a control scheme be implemented even in principle if nodes represent species populations and a control signal calls for a fractional change in a node state variable over the course of minutes? Furthermore, how do the authors interpret the methods by which cycles are controlled in linear control theory (see e.g. doi 10.1126/science.1242063 and its supplementary material) in the context of ecological management? How frequent are cycles in the authors' ecological networks? I should emphasize that I am a fan of control theory and I think it is interesting to apply it as the authors do in this ms. However, the caveats that come along with the points I outline above need to be made clear for the reader. The authors might be interested in doi 10.1063/1.4931570 for a discussion of some relevant ideas.

    * Acepted; changed.*

@@

@@

@@

Second, a node with no outgoing edges will clearly never be a superior node and will therefore not be a driver node. I assume that this never happens in the mutualistic networks under consideration, but this probably bears mentioning so as to help develop the reader's intuition for this assessment. In any event it would be useful for the reader to see a simple network (or several) to provide clear examples of driver nodes and non-driver nodes. For instance, how can a node with an outgoing edge not be a driver node?

Finally, the term "driver species" (or more generically, "driver node") has a specific meaning in the control literature (doi 10.1038/nature10011): what the authors here denote with n_d on line 118. To avoid confusion with those familiar with the control theory literature (and those readers who are inspired by this ms to delve into it), I suggest a change of terminology. Perhaps "key species"?

@@

@@

    *Acepted; changed. The sentence now reads "To bridge this gap, we propose methods to apply previous theory on the control of complex systems into an ecological context and implement them in empirical data"*

@@

@@

@@

Only one type of interaction is being considered here when species are involved in multiple types of interactions. An assumption that is not explicit in the text is that all the factors controlling population densities are the interactions between plants and pollinators. The boundaries of a network, including the types of interactions that are considered or not, are often arbitrarily defined. However, this approach is highly dependent on the topology of the network and thus on how we define the network. I know this limitation applies to any study on system dynamics, but it becomes more serious when we argue that management decisions can be based on a single technique. I have two suggestions. First, the authors should discuss in detail these assumptions and their limitations. Second, It would be helpful if the authors could test the robustness of their results to sampling. The results show that invasive species change manageability. This suggests that manageability may be sensitive to sampling. A sensitivity analysis where species are added and removed from the network under different wiring schemes would help us understand the behavior of the manageability index when networks are undersampled and would allow a more informed discussion on how the choices we make when delimiting the network affect estimates of controllability. Weak links should not have a strong effect on controllability and considering the networks were built using visitation frequency, most of the strong interactions were likely sampled. Still, it would be great to show that the method is robust and this would make it easier to justify that even in the absence of complete sampling this technique is in fact useful for management.

@@

Ln 95-98. It is very hard to keep track of all those definitions - matching, matched nodes, matched links, matching size, matching weight, maximum matching. I found myself going back and forth several times between the framework, methods and results sections. A glossary would help a lot.

Ln 99. Indicate how the matching is represented in the figure. e.g.,  "(dark grey links in Figure 1a)"

Figure 1.Indicate the driver nodes in these networks

Ln 128. Be clear about the meaning of resources here. I understand you mean financial resources not  resources in the ecological sense.

>>>>

---
title: "Response to reviewer comments ELE-00738-2016"
date: "21 November 2016"
header-includes:
   - \usepackage{setspace}
output: pdf_document
classoption: a4paper
bibliography: references.bib
---

\onehalfspacing

@@

We are grateful for the opportunity to resubmit a revised manuscript for your consideration. We have endeavored to incorporate all reviewers' feedback into the manuscript and we feel the attached draft constitutes an important improvement from the previous version. We were impressed by the quality of the reviews; they all were thoughtful, clear, and constructive. It is evident that all three anonymous reviewers have given a fair and careful consideration to our manuscript. We would like to thank them for the truly valuable feedback.

The reviewers identified major limitations in our manuscript. Specifically with regards to the terminology, the underlying  structural control, and the lack of detail in the presentation of the empirical networks. In the current version, we have thoroughly addressed all these concerns and improved both the robustness of the analyisis and the clarity of the concepts. Detailed responses to each comment, one by one, are provided below.

@@

    *We revised the text to better reflect the contribution of @Tylianakis2010 at the same time we added a reference to a more recent paper by @Kaiser-Bunbury2015 who explicitly deals with oportunities, limitations and implications of applying network metrics in conservation science.*

- Lines 95-97: "The number of nodes necessary to fully control a complex network can be calculated by counting the number of unmatched nodes in the network's maximum matching (Liu et al. 97 2011)." This statement must be contextualized by the assumed nature of the dynamical relationships in this framework. It therefore seems to me that this statement should be preceded by a discussion of those details (indeed, the authors argue that this is necessary in the Introduction, on lines 48-49). Furthermore, the authors need to discuss the appropriateness of this framework to ecological systems. The input control signals that drive even relatively simple networks to specified states can be very sensitive to initial conditions and complicated (indeed, they usually aren't calculated at all). How can a control scheme be implemented even in principle if nodes represent species populations and a control signal calls for a fractional change in a node state variable over the course of minutes?

    *Thanks to this comment and those from other reviewers have helped us identify some shortcomming on the way we were introducing the control framework in the paper. We have largely rewritten the 'Theoretical framework' section and we start by offering a brief overview of control theory which emphasize on a dynamical perspective to complex networks (Line XXX). We then follow with a paragraph devoted to how structural controlability is unable to provide information about the particulars of the control signal and the feasibility of its application (Line XXX). Without disregard the usefulness of the proposed approach, we also mention in this paragraph, and later on the discussion (Line XXX), how recently developed numerical methods [@Cornelius2013] might offer exiting results in this regard.*

- Furthermore, how do the authors interpret the methods by which cycles are controlled in linear control theory (see e.g. doi 10.1126/science.1242063 and its supplementary material) in the context of ecological management? How frequent are cycles in the authors' ecological networks? I should emphasize that I am a fan of control theory and I think it is interesting to apply it as the authors do in this ms. However, the caveats that come along with the points I outline above need to be made clear for the reader. The authors might be interested in doi 10.1063/1.4931570 for a discussion of some relevant ideas.

    *Thanks for pointing this out. In fact we thoroughly implememnted the approach of [@Ruths2014] at initial stages of the project. However we found that the profile of the network depends to an extreme degree on the chosen link directionality and very little on the particulars of a given network. Because of that reason we decided to exclude this approach altogether from the final draft. Nevertheless, due to this comment and one of Referee 2 we included this treatment in the Supplementary Materials Section XX and reference it briefly in the manuscript (Line XXX). In SM SectionXX we explain that because we focus on bipartite networks, cycles only appear, and in fact dominate, when we consider the case of bidirectional links. We are particularly grateful for the reference to @Motter2015. We were unaware of this paper and has been extremely useful to contextualise our work in the broader control theory.*

@@

    *Indeed most of the literature treats the problem of controlability on unweighted networks. It has been shown, however, at least in pollination networks like the ones used in our paper, that the interaction weights carry important information about the patterns and processes in the community. In the paper we use the visitation weights directly or indirectly in two different aspects. First, we use them to determine the direction of control (via the dependence asymetry). Second, which directly adresses the reviewer's comment, we use it to give priority to matchings that have the largest weight, which potentially also have the largest impact on the network state for a given management intervention. Because the weights determine to some extent the dynamical relationships between species, our approach attempt to incorporate more realism to the structural controllability approach. We acknoweledge our presentation was confusing, especially for readers familiar wit the control literature, and therefore have added a whole paragraph in the method describing our rationale (Lines XXX)*

@@

    *We have rewritten that section and the equivalent sentence now reads "It has been shown that a particular network that can be expressed as a state space system of the form $\dot{x} = Ax + Bu(t)$ is controllable if there exists a minimum number of nodes in the network (N~D~ > 0) to which is possible to aply an external control signal that propagates trough the network to in turn drive the state of every other node in the network to a particular configuration [@Kalman1963; @Liu2011]."*

@@

    *We are sorry for the impresicion. It should have been "all possible maximum cardinality matchings". In any case, because of the suggestion about the driver species concept we have decided to remove this sentence. We also reworded the text in Lines XXX to clarify why our approach to identify species importance is not exclsively centered on maximum matchings.*

- Second, a node with no outgoing edges will clearly never be a superior node and will therefore not be a driver node. I assume that this never happens in the mutualistic networks under consideration, but this probably bears mentioning so as to help develop the reader's intuition for this assessment. In any event it would be useful for the reader to see a simple network (or several) to provide clear examples of driver nodes and non-driver nodes. For instance, how can a node with an outgoing edge not be a driver node?

    *Thanks to this suggestion and those of other reviewers, we have made important changes to Figure 1. On it we include a better ilustration of our methodology for a simple network while paying special effort to bring clarity to the different node categories as superior, matched and unmatched nodes. TO DO*

- Finally, the term "driver species" (or more generically, "driver node") has a specific meaning in the control literature (doi 10.1038/nature10011): what the authors here denote with n_d on line 118. To avoid confusion with those familiar with the control theory literature (and those readers who are inspired by this ms to delve into it), I suggest a change of terminology. Perhaps "key species"?

    *Overall our choice of terminology was poor. We have eliminated the term "driver species" from the manuscript and now just mention, in the referee's words, how the frequency with which species exist on key paths trough the network provides indication of their relative importance.*

@@

    *Thanks for pointing this out. As suggested, we have included a figure that help explains our method to compute all possible maximum cardinality matchings. However because the methodology itself is not one of our main contributions and because of space limitations we have included that figure in the Supplementary Material. TO DO.*

@@

    *The sentence now reads "To bridge this gap, we propose methods to apply previous theory on the control of complex systems into an ecological context and implement them in empirical data"*

@@

    *We have pretty much rewritten the "Theoretical framework" section following the reviewers advise. We start with a brief introduction to structural controllability (Lines XXX), follow with a definition of manageability (Lines XXX) and finalise with details and computational considerations for its calculation.*

@@

     *Thanks for the observation. We have reworded large segments of this section. Specifically now explain how the approach builds on existing structural controllability theory by incorporating the weight of the interactions. Moreover, now we included more details about the rationale behind our approach. By being more specific we hope the reader to be able to identify the theretic and computational novelties of our approach with respect to existing "ranking" methods. We also strive to further clarify the novelty in the discussion section (Lines XXX). TO DO*

@@

 - Only one type of interaction is being considered here when species are involved in multiple types of interactions. An assumption that is not explicit in the text is that all the factors controlling population densities are the interactions between plants and pollinators. The boundaries of a network, including the types of interactions that are considered or not, are often arbitrarily defined. However, this approach is highly dependent on the topology of the network and thus on how we define the network. I know this limitation applies to any study on system dynamics, but it becomes more serious when we argue that management decisions can be based on a single technique. I have two suggestions. First, the authors should discuss in detail these assumptions and their limitations.

    *PARTIAL ANSWER: We have included some comments about the general limitations of the structural controlability approach in the "Theoretical Framework" section (Lines XXX). Furthermore*

Second, It would be helpful if the authors could test the robustness of their results to sampling. The results show that invasive species change manageability. This suggests that manageability may be sensitive to sampling. A sensitivity analysis where species are added and removed from the network under different wiring schemes would help us understand the behavior of the manageability index when networks are undersampled and would allow a more informed discussion on how the choices we make when delimiting the network affect estimates of controllability. Weak links should not have a strong effect on controllability and considering the networks were built using visitation frequency, most of the strong interactions were likely sampled. Still, it would be great to show that the method is robust and this would make it easier to justify that even in the absence of complete sampling this technique is in fact useful for management.

@@

- Ln 95-98. It is very hard to keep track of all those definitions - matching, matched nodes, matched links, matching size, matching weight, maximum matching. I found myself going back and forth several times between the framework, methods and results sections. A glossary would help a lot.

    *Thanks. We have rewritten a large section of the methods and tried to be more clear when defining new terms by providing some examples and using a more logical sequence to go from one concept to the next (Lines XXX). We have also included a short glossary with the most important definitions. TO DO*

- Ln 99. Indicate how the matching is represented in the figure. e.g.,  "(dark grey links in Figure 1a)". Figure 1.Indicate the driver nodes in these networks

    *We have redesigned Figure 1 and now we indicate more clearly how the matching works and specify which are the superior, matched, and unmatched nodes. Also because we also illustrate the methods to obtain the relative importance, it is easier to highlight which are the key species in a simple network.*

- Ln 128. Be clear about the meaning of resources here. I understand you mean financial resources not  resources in the ecological sense.

    *Indeed. We apologise for the imprecision. We indeed mean financial resources. We have modified the sentence to improve clarity and it now reads "Ecologically, these distinctions are relevant because management and conservation resources are limited and therefore ecological interventions should ideally be focused on species that might provide the largest impact."*
